---
title: "Applied DRA with sdcMicro Draft"
author: "Shaina Trevino"
date: "1/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sdcMicro)
```

## Import Data

Chose variables, clean data (filter year, aggregate schools), random subsample, simulate data, export for use

randomly simulated, similar descriptives/distributions, correlations are not maintained. 

```{r import}
sim_df <- rio::import(here::here("data", "sim_df.csv")) %>% 
  mutate_if(is.character, as.factor)
#need to deal with missing data for characters/factors that are blank now

summary(sim_df)
psych::describe(sim_df)
```

## Understanding Data and Context

## Selecting Key Variables

- different scenarios
- run many times with different keys and see which variables are leading to most risk (e.g., most suppression, most unique keys)

## Assessing Risk

Make sure there are similar unique cases/riskiness in simulated data

###Categorical Variables

#### Sample Frequencies/Uniqueness

move info from below here

```{r samp-freq}
#vector of categorical variable names
catvars <- sim_df %>% 
  select_if(is.factor) %>% 
  colnames()
#dataset with only categorical variables
catdf <- sim_df %>% 
  select_if(is.factor)

#sdcMicro function
freq_keys <- freqCalc(sim_df, keyVars = catvars) #creates sdcMicro object


#number of unique rows
freq_keys$n1
#number of rows with freq = 2
freq_keys$n2
#frequency counts for each row
counts <- freq_keys$fk
#combine dataframe with frequencies counts for each row
freq_counts <- cbind(sim_df, counts)
head(freq_counts[, c(colnames(sim_df), "counts")])

#aggregate to get frequencies for each key/combination
agg_counts <- aggregate(counts ~ ., catdf, mean)
nrow(agg_counts) #number of possible keys/combinations

sum(agg_counts$counts == 1) #unique keys/combinations

sum(agg_counts$counts == 2) #keys/combinations with freq = 2

##above is same info in different ways (maybe dont need aggregate unless youw ant to know number of possible keys - can get another way I am sure)

```


#### K-anon violations
how is it different from above (switch section)
```{r k-anon-vio}
#vio k-anon violations
freq_keys
```


#### Individual Risk

```{r ind-risk}


##INDIVIDUAL RISK CALCULATION 
indivf <- indivRisk(freq_keys)
inriskvec <- indivf$rk

freq_risk <- cbind(freq_counts, inriskvec) #- higher is worse. 1 = unique? indiv risk based on counts 15 = .066 risk

```

#### Global Risk

f(k) frequency count of each key

```{r first-risk}
#need sdc object
#fk <- 
```

